---
title: "Trial-estimated parameters for the SALT cost-effectiveness model"
subtitle: "Replication of Bernabe-Ortiz et al. (2020), and new mortality and CVD estimates"
author: "Darwin Del Castillo"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    number-sections: true
    theme: readable
    df-print: kable
    code-fold: true
editor_options:
  chunk_output_type: console
---

```{r setup}
#| include: false
here::i_am("salt_results.qmd")
library(data.table)
library(ggplot2)
library(survival)
res <- readRDS(here::here("output", "params_est.rds"))
fmt_ci <- function(e, l, h, d = 2) {
  # a separated model returns an astronomical point estimate and an infinite
  # bound; print those as "not estimable" rather than 818784100.00 (Inf, Inf)
  ok <- is.finite(e) & is.finite(l) & is.finite(h) & abs(e) < 1e3
  out <- rep("not estimable", length(e))
  out[ok] <- sprintf(paste0("%.", d, "f (%.", d, "f, %.", d, "f)"),
                     e[ok], l[ok], h[ok])
  out
}
```

# Why this document exists

The microsimulation on `exp/globorisk` takes CVD risk from the Globorisk office
equation, which is external to this trial, validated only for ages 40–74, and
carries an unresolved 3–4× level disagreement with GBD Peru incidence. The trial
has seven waves of follow-up on `r res$long_n$ids` people and can estimate some
of these quantities itself.

Bernabe-Ortiz et al. (*Nature Medicine* 2020;26:374–8) analysed this same
dataset, and the supplementary material gives the exact Stata specification.
That makes replication a **validation gate**: if `clean_long()` reproduces the
published numbers, the cleaning is trustworthy, and the same machinery can be
pointed at the endpoints the paper never analysed — death and CVD.

Everything below is produced by `run_params_est.R`, which caches to
`output/params_est.rds`. This document only renders that file.

# Replication

## Cohort

```{r cohort}
#| echo: false
data.table(
  Quantity = c("Person-waves", "Individuals", "Households (codhogar)"),
  Reproduced = c(res$long_n$rows, res$long_n$ids, res$long_n$households),
  Published  = c("16,632 samples", "2,376", "not reported")
)
```

## Wide-vs-long event reconciliation

The trial data exist in two independently structured files. The long file is one
row per person-wave; the wide file is one row per person with a column block per
visit. Both should encode the same events. They do:

```{r recon}
#| echo: false
res$wide_events$recon
```

Zero mismatches on every event variable, so `clean_long()` is not dropping or
duplicating events during the wave assignment.

## Intervention effects on blood pressure and hypertension incidence

Published values are Table 2 (blood pressure) and Supplementary Table 3
(hypertension incidence). "Match" means the point estimate is within 0.10 of the
published one, on the mmHg scale for blood pressure and the hazard-ratio scale
for incidence.

```{r replication}
#| echo: false
r <- copy(res$replication)
r[, .(Quantity = quantity, Source = source,
      Published = fmt_ci(published, published_low, published_high),
      Reproduced = fmt_ci(estimate, low, high, 3),
      Difference = round(diff, 3),
      Verdict = verdict)]
```

All four blood-pressure coefficients reproduce to two decimals. The hypertension
hazard ratios reproduce to within 0.06.

### One quantity that does not reproduce: person-time

```{r pt-residual}
#| echo: false
pt <- res$rates[endpoint == "htn_composite"]
data.table(
  Quantity = c("Incident cases, control", "Incident cases, intervention",
               "Person-years, control", "Person-years, intervention"),
  Published = c(107, 133, 1961.1, 2712.3),
  Reproduced = c(pt[arm == "Control", events], pt[arm == "Intervention", events],
                 round(pt[arm == "Control", py], 1),
                 round(pt[arm == "Intervention", py], 1))
)[, Difference := sprintf("%+.1f%%", 100 * (Reproduced - Published) / Published)][]
```

The **event counts are exact** — 107 and 133, matching Supplementary Table 3
person for person. Person-time runs about 4% high in both arms.

Because the excess is nearly proportional across arms it cancels in the hazard
ratio, which is why the HR still reproduces. The likely cause is the censoring
convention for the final interval: this implementation carries each person to
their last dated visit, whereas the published analysis reports 1,891 individuals
against the 1,914 eligible, implying 23 were dropped for having no usable
follow-up. That difference is not recoverable from the published description. It
is recorded here rather than tuned away.

## Two specification choices that had to be recovered

Neither is stated unambiguously in the paper, and both change the answer
materially. They are recorded here because the published text alone does not
determine them.

### The survival time axis is time-on-study, not calendar time

The methods say "Cox's proportional hazard modeling on a calendar time axis, to
account for time trends". Empirically that is not what produced the published
number:

```{r axis}
#| echo: false
res$axis_check[, .(`Time axis` = axis,
                   `HR (95% CI)` = fmt_ci(hr, low, high, 3),
                   `Published` = "0.49 (0.34, 0.71)")]
```

Time since each person's own baseline visit reproduces the published hazard
ratio; calendar date does not. Stata's `dtime` was therefore time on study.
`build_survival_set()` returns both axes so this stays testable.

### Covariate adjustment uses baseline values, not time-varying ones

The paper adjusts for "age, sex, education, wealth index and BMI **at
baseline**". Using the time-varying columns instead moves the fully adjusted
hypertension HR from 0.489 to 0.586 — outside the published interval. The
`baseline_covars()` helper carries each person's wave-0 value to all their rows.

### Inference is CR0 with a normal critical value

Model-based standard errors are roughly half the published width (0.25 vs 0.40
for SBP), because they ignore village clustering. Stata's
`vce(cluster codvilla)` is a CR0 sandwich reported against a normal critical
value, which is what `fit_bp_model()` uses. CR2 would be the better small-sample
choice with only six clusters, but it is not computable here: it requires a
matrix square root per cluster and these clusters hold ~2,500 observations each.
CR1 and CR3 standard errors are carried in the output for comparison.

```{r se-compare}
#| echo: false
res$bp[, .(Outcome = outcome, Adjustment = adjust,
           Estimate = round(estimate, 3),
           `SE model-based` = round(se_model, 3),
           `SE CR0` = round(se, 3),
           `SE CR1` = round(se_cr1, 3),
           `SE CR3` = round(se_cr3, 3))]
```

## Proportional hazards

```{r schoenfeld}
#| echo: false
res$schoenfeld[, .(`Schoenfeld chi-squared` = round(chisq, 3),
                   `p` = round(p, 3),
                   `Published p` = 0.40)]
```

# New endpoints

The paper did not analyse mortality or CVD. Both are estimated here with the
same design — shared gamma frailty on the village, time-on-study axis.

## Event counts and crude rates

```{r rates}
#| echo: false
res$rates[, .(Endpoint = endpoint, Arm = arm, Events = events,
              `Person-years` = round(py, 1),
              `Rate per 100 py` = round(rate_100py, 2))]
```

## Hazard ratios

```{r hr}
#| echo: false
res$incidence[, .(Endpoint = label, Adjustment = adjust, Events = events,
                  `HR (95% CI)` = fmt_ci(hr, low, high, 3),
                  `HR (frailtyEM)` = round(hr_emfrail, 3),
                  Converged = is.na(warning))]
```

## Mortality

`r res$rates[endpoint == "death", sum(events)]` deaths over
`r round(res$rates[endpoint == "death", sum(py)])` person-years. This is the one
new endpoint that is usable.

```{r km}
#| echo: false
#| fig-height: 4
d <- res$sets$death
km <- survfit(Surv(g0, g1, event) ~ intervencion, data = d)
plot(km, col = c("#2b6cb0", "#b7791f"), lwd = 2, conf.int = TRUE,
     xlab = "Years since baseline visit", ylab = "Survival",
     ylim = c(0.95, 1))
legend("bottomleft", c("Control", "Intervention"),
       col = c("#2b6cb0", "#b7791f"), lwd = 2, bty = "n")
```

The time-and-cluster-adjusted estimate is the one to use. The fully adjusted
model fits five covariates to 34 events — roughly seven events per parameter,
well below the conventional ten — so its interval is not trustworthy even though
it is narrower.

```{r death-hr}
#| echo: false
res$incidence[label == "death",
  .(Adjustment = adjust, Events = events,
    `HR (95% CI)` = fmt_ci(hr, low, high, 3),
    `Free parameters` = fifelse(adjust == "full", 6L, 1L),
    `Events per parameter` = round(events / fifelse(adjust == "full", 6, 1), 1))]
```

## CVD: why these estimates should not be used

The flag-based CVD estimate is large, positive and nominally in the wrong
direction. It is an artefact of how the CVD history variables were collected,
and three pieces of evidence establish that.

### 1. The flags reverse themselves

`derrame`, `infarto`, `insuficiencia` and `otracor` are point-in-time
self-reports, not a cumulative history. Nearly every reported "Yes" is followed
by a "No" at the next visit:

```{r revert}
#| echo: false
res$wide_events$revert[, .(Variable = var, `Ever reported` = ever,
                           `At baseline` = prevalent,
                           `First after baseline` = first_after_w0,
                           `Reverse later` = reverts)]
```

A variable that contradicts itself at the next visit cannot distinguish a new
event from a re-report of an old one.

### 2. Only stroke carries a corroborating event date

The wide file keeps date fields the long file drops, and only for `derrame`.
At baseline, `derrame_a` records years since the stroke — every baseline case is
historical, some by decades:

```{r stroke-prev}
#| echo: false
res$wide_events$stroke$prevalent[, .(codigo, `Years since stroke` = years_since,
                                     `Doctor-confirmed` = confirmed)]
```

At follow-up, exactly two people report a stroke, and both carry a month and
year consistent with the wave they reported it in:

```{r stroke-inc}
#| echo: false
res$wide_events$stroke$incident[, .(codigo, Wave = wave,
                                    `Event date` = format(event_date, "%b %Y"),
                                    `Doctor-confirmed` = confirmed)]
```

`infarto`, `insuficiencia` and `otracor` have no date fields in either file, so
none of their "incident" cases can be corroborated at all.

### 3. The verifiable count is 2, not 13

```{r cvd-compare}
#| echo: false
res$incidence[label %in% c("cvd_flag", "cvd_verified") & adjust == "time_cluster",
  .(Definition = c("Any flag turns Yes after baseline",
                   "Stroke with a corroborating event date")[
                     match(label, c("cvd_flag", "cvd_verified"))],
    Events = events,
    `HR (95% CI)` = fmt_ci(hr, low, high, 2))]
```

With two events, both in intervention person-time and none in control, the model
is separated and the hazard ratio is not defined. That is the honest ceiling on
what this trial can say about CVD incidence.

**Recommendation: Globorisk remains the CVD risk source.** The trial cannot
estimate a CVD hazard, and the flag-based estimate should not be carried into
the microsimulation under any adjustment.

# What this contributes to the microsimulation

```{r contrib}
#| echo: false
data.table(
  Parameter = c("delta_sbp", "delta_dbp", "Hypertension incidence",
                "All-cause mortality", "CVD-cause mortality",
                "CVD incidence", "Costs", "Utilities / disability weights"),
  Source = c("This analysis", "This analysis", "This analysis",
             "This analysis", "This analysis (11 deaths)",
             "Globorisk — trial cannot estimate it",
             "Costing study", "GBD / placeholder"),
  Status = c("replicated", "replicated", "replicated",
             "new, usable", "new, sparse",
             "unchanged", "unchanged", "unchanged")
)
```

The blood-pressure effect the microsimulation already uses
(`delta_sbp = -1.29`) is confirmed by this replication. The new contribution is
mortality: `r res$rates[endpoint == "death", sum(events)]` deaths estimated with
the trial's own design, replacing pooled counts that ignored village clustering.

# Session

```{r session}
#| echo: false
data.table(
  Package = c("R", "lme4", "clubSandwich", "survival", "frailtyEM", "data.table"),
  Version = c(paste(R.version$major, R.version$minor, sep = "."),
              as.character(packageVersion("lme4")),
              as.character(packageVersion("clubSandwich")),
              as.character(packageVersion("survival")),
              as.character(packageVersion("frailtyEM")),
              as.character(packageVersion("data.table")))
)
```
