---
title: "Cost-effectiveness analysis of a salt substitute intervention in Northern Peru"
author: "Darwin Del Castillo"
date: today
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
here::i_am("salt_results.qmd")
source(here::here("R", "cea.R"))
source(here::here("R", "psa.R"))
source(here::here("R", "owsa.R"))
results <- readRDS(here::here("output", "cea_results.rds"))
```

# Main results

## Table 1. Base-case parameters

```{r tab1}
#| echo: false
#| include: true
#| message: false

p <- results$params
knitr::kable(data.frame(
  Parameter = c("Cycles", "Cycle length (yr)", "Discount rate",
                "Treatment effect on SBP (mmHg)"),
  Value     = c(as.character(as.integer(p$n_t)),
                format(round(p$cycle_length, 3), nsmall = 3),
                paste0(p$d_c * 100, "%"), paste0(p$delta_sbp, " mmHg"))
))
```

## Table 2. Outcomes and definitions

```{r tab2}
#| echo: false
#| include: true
#| message: false
knitr::kable(results$cea$summary, digits = 3)
```

## Figure 1. Markov model structure

![Model structure](figs/model_structure.svg)

## Figure 2. Cost-effectiveness plane

```{r fig2}
#| echo: false
#| include: true
#| message: false
ce_plane(results$cea)
```

## Figure 3. Cost-effectiveness acceptability curve

```{r fig3}
#| echo: false
#| include: true
#| message: false
if (is.null(results$psa)) {
  cat("Not run. Re-run with `SALT_PSA_N=1000 Rscript run_microsim.R`.")
} else {
  ceac_plot(ceac(results$psa))
}
```

# Sensitivity analyses

## Figure 4. One-way sensitivity analysis (tornado)

```{r fig4}
#| echo: false
#| include: true
#| message: false
if (is.null(results$owsa)) {
  cat("Not run. Re-run with `SALT_OWSA=TRUE Rscript run_microsim.R`.")
} else {
  tornado_plot(results$owsa)
}
```

## Table 3. One-way sensitivity analysis ranges

```{r tab3}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$owsa)) {
  knitr::kable(results$owsa[, .(Parameter = label, Low = low, High = high,
                                `ICER at low` = out_low,
                                `ICER at high` = out_high,
                                Spread = spread)], digits = 2)
}
```

## Figure 5. Cost-effectiveness plane (PSA)

```{r fig5}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$psa)) ce_plane_psa(results$psa)
```

## Table 4. Probabilistic results

```{r tab4}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$psa)) {
  knitr::kable(psa_summary(results$psa), digits = 3)
}
```

## Figure 6. PSA convergence

```{r fig6}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$psa)) psa_convergence(results$psa, "icer_daly")
```