---
title: "Cost-effectiveness analysis of a salt substitute intervention in Northern Peru"
author: "Darwin Del Castillo"
date: today
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
here::i_am("salt_results.qmd")
source(here::here("R", "cea.R"))
source(here::here("R", "psa.R"))
source(here::here("R", "owsa.R"))
results <- readRDS(here::here("output", "cea_results.rds"))
```

# Main results

## Table 1. Base-case parameters

```{r tab1}
#| echo: false
#| include: true
#| message: false

p <- results$params
knitr::kable(data.frame(
  Parameter = c("Horizon", "Cycles (total)", "Cycles observed in trial",
                "Cycle length (yr)", "Discount rate",
                "Treatment effect on SBP (mmHg)",
                "Post-trial exposure", "Post-CVD excess mortality hazard"),
  Value     = c(sprintf("%.1f yr (to age %d)", p$n_t * p$cycle_length, p$max_age),
                as.character(as.integer(p$n_t)),
                as.character(as.integer(p$n_trial)),
                format(round(p$cycle_length, 3), nsmall = 3),
                paste0(p$d_c * 100, "%"), paste0(p$delta_sbp, " mmHg"),
                p$post_trial, format(p$hr_postcvd))
))
```

The first `r as.integer(p$n_trial)` cycles follow the trial's own measurement
rhythm, one cycle per inter-visit interval. Beyond them each participant's last
observed blood pressure is carried forward, so every post-trial difference
between arms is the treatment effect itself rather than a projection artefact.
Age still advances, so absolute risk continues to climb.

## Table 2a. Post-trial exposure scenarios

```{r tab2a}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$cea_stop)) {
  sc <- function(x, lab) data.frame(
    Scenario           = lab,
    `Incremental cost` = x$d_cost,
    `DALYs averted`    = x$daly_averted,
    `QALYs gained`     = x$qaly_gained,
    `ICER per DALY`    = x$icer_daly,
    check.names = FALSE)
  knitr::kable(rbind(
    sc(results$cea,      "Exposure continues (base case)"),
    sc(results$cea_stop, "Exposure stops at trial end")), digits = c(0, 2, 6, 6, 1))
}
```

Under the base case both the salt-substitute cost and its effect on blood
pressure persist for the whole horizon. The sensitivity scenario ends both when
the trial does: the arms' blood pressures converge at the first projected cycle
and no further intervention cost accrues, while events already averted during
the trial stay averted.

## Table 2. Outcomes by arm

```{r tab2}
#| echo: false
#| include: true
#| message: false
knitr::kable(results$cea$summary, digits = 3)
```

## Figure 1. Markov model structure

![Model structure](figs/model_structure.svg)

## Figure 2. Cost-effectiveness plane

```{r fig2}
#| echo: false
#| include: true
#| message: false
ce_plane(results$cea)
```

## Figure 3. Cost-effectiveness acceptability curve

```{r fig3}
#| echo: false
#| include: true
#| message: false
if (is.null(results$psa)) {
  cat("Not run. Re-run with `SALT_PSA_N=1000 Rscript run_microsim.R`.")
} else {
  ceac_plot(ceac(results$psa))
}
```

# Sensitivity analyses

## Figure 4. One-way sensitivity analysis (tornado)

```{r fig4}
#| echo: false
#| include: true
#| message: false
if (is.null(results$owsa)) {
  cat("Not run. Re-run with `SALT_OWSA=TRUE Rscript run_microsim.R`.")
} else {
  tornado_plot(results$owsa)
}
```

## Table 3. One-way sensitivity analysis ranges

```{r tab3}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$owsa)) {
  knitr::kable(results$owsa[, .(Parameter = label, Low = low, High = high,
                                `ICER at low` = out_low,
                                `ICER at high` = out_high,
                                Spread = spread)], digits = 2)
}
```

## Figure 5. Cost-effectiveness plane (PSA)

```{r fig5}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$psa)) ce_plane_psa(results$psa)
```

## Table 4. Probabilistic results

```{r tab4}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$psa)) {
  knitr::kable(psa_summary(results$psa), digits = 3)
}
```

## Figure 6. PSA convergence

```{r fig6}
#| echo: false
#| include: true
#| message: false
if (!is.null(results$psa)) psa_convergence(results$psa, "icer_daly")
```